---
title: "SPPH 581c - Assignment 4"
author: "Patrick Daniele"
date: "February 20, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<p>I'm going to complete this as an R-Markdown document as some of the packages I use require R-Studio and a very up to date version of R.</p>


<p>Load Libraries </p>
```{r}
library(survival)
library(tidyverse)
library(survminer)
library(ggfortify)
library(flexsurv)
```
<p>Incase you haven't seen it before, the "tidyverse" is a set of packages that makes using R much better (In my opinion).</p>


<p>Import the data and basic tidying</p>
```{r}
temp <- read.csv('C:/Users/pdaniele/Desktop/SPPH 581c/Assignment 4/alzheimers.csv', header=T)
str(temp)

#Note to self, dates are factors. I don't think we need them, so I won't do anything about them now.

#Change drug to a factor
temp2 <- temp %>% 
  mutate(drug=factor(drug))

# Change it to a tibble for easier processing
data <- as_data_frame(temp2)
data
```


<p> Basic data exploration</p>
```{r}
## Checking normality of Age.
hist(data$age)
## Looks good to me

#Checking how many had events
table(data$censor)

## Checking How many on each drug
table(data$drug)
## Balanced!

## Check the association between time and censor, and age as per Jason's recommendation
fit1 <- lm(time ~ age, data=data)
summary(fit1)
## Pretty strong negative association. Makes sense as older people would die younger.
fit2 <- glm(censor~age, data=data)
summary(fit2)

## Not sig. Good! Wouldn't want to see any differences in people who are censored. 
## This helps to validate our assumption of non-informative censoring.
fit3 <- glm(censor~drug, data=data)
summary(fit3)
## Same for Drug.

```

<p> Three majors sections:</p>

<ol>
  <li>Non-Parametric: Kaplan Meier Estimates</li>
  <li>Semi-Parametric: Cox Proportional Hazards Model</li>
  <li>Fully Parametric: Weibull, Exponential, Log-Normal, Log-Logistic</li>
</ol>


<h1>Non-Parametric</h1>
```{r}
## Overall
fit_KM <- survfit(Surv(time, censor) ~ 1, data)

fit_KM %>% 
  summary()

## Try different survival plot methods
#Extensions of GGPlot2

fit_KM %>% 
  ggsurvplot(risk.table = TRUE, ggtheme=theme_classic())

fit_KM %>% 
  autoplot(conf.int=F)

## By Drug
fit_KM_drug <- survfit(Surv(time, censor) ~ drug, data)

fit_KM_drug %>% 
  summary()

survdiff(Surv(time, censor) ~ drug, data)

## Try different survival plot methods
fit_KM_drug %>% 
  autoplot(conf.int=F)

fit_KM_drug %>% 
  ggsurvplot(risk.table = TRUE, ggtheme=theme_classic())


## Let's get both the by drug and overall on the same plot
plot(fit_KM, conf.int=F, xlab='Time (Months)', ylab='Survival', lwd=2)
lines(fit_KM_drug, lty=c(2,2), lwd=c(2,2), col=c("#8E8E8E","#5CB2C8"))
legend('topright', c('Drug 0', 'Drug 1'), lwd=2, lty=2, col=c("#8E8E8E","#5CB2C8"))
text(40, 1, 'Log-rank p=0.000971')

## Side by Side KM Plots
plot(fit_KM_drug, conf.int=T, xlab='Time (Months)', ylab='Survival', lwd=2, col=c("#8E8E8E","#5CB2C8"))
legend('topright', c('Drug 0', 'Drug 1'), lwd=2, lty=2, col=c("#8E8E8E","#5CB2C8"))

```
<p> This is completely descriptive, so I think I would always do this regardless of what modeling technique I chose in the end</p>


<h1>Semi-Parametric</h1>
```{r}

## Without Age
coxfit1 <- coxph(Surv(time, censor) ~ drug, data)
coxfit1 %>% 
  summary()

## With Age
coxfit2 <- coxph(Surv(time, censor) ~ drug + age, data)
coxfit2 %>% 
  summary()

## Check model Fit and Assumptions (PH)
coxfit1 %>% 
    AIC()

coxfit1 %>% 
  cox.zph() %>% 
  plot()

coxfit2 %>% 
    AIC()

coxfit2 %>% 
  cox.zph() %>% 
  plot()

plot(residuals(coxfit2, type='martingale'))
plot(residuals(coxfit2, type='schoenfeld'))

#Plotting the Cumulative Hazard
cumhaz.plot <- survfit(Surv(time, censor) ~ drug, data, type='fleming') 

plot(cumhaz.plot, lty=2:3, fun='cumhaz', xlab='Time (Months)', ylab='Cumulative Hazard')
## Global test looks good. Plots look OK to me. 
## Residuals look good.
## Age improves fit.

## Plotting survival for the drug=1 group.
autoplot(survfit(coxfit1, newdata=data.frame(drug=1)), conf.int=F)

##Survival for Age = 50
autoplot(survfit(coxfit2, newdata=data.frame(drug=0, age=50)), conf.int=F)
autoplot(survfit(coxfit2, newdata=data.frame(drug=1, age=50)), conf.int=F)

```
<p> Assumptions look good, fit is reasonable. I'll compare it to the parametric curves at the end</p>


<h1>Fully-Parametric</h1>
```{r}
#Exponential Distribution
fit_AFT_exp <- survreg(Surv(time, censor) ~ drug + age, data, dist='exponential')
fit_AFT_exp %>% 
  summary()

plot(residuals(fit_AFT_exp, type="deviance"))

fit_AFT_exp %>% 
    AIC()

#Weibull Distribution
fit_AFT_W <- survreg(Surv(time, censor) ~ drug + age, data, dist='weibull')
fit_AFT_W %>% 
  summary()

fit_AFT_W %>% 
    AIC()

plot(residuals(fit_AFT_W, type="deviance"))

#Lognormal
fit_AFT_lognorm <- survreg(Surv(time, censor) ~ drug + age, data, dist='lognormal')
fit_AFT_lognorm %>% 
  summary()

fit_AFT_lognorm %>% 
    AIC()

plot(residuals(fit_AFT_lognorm, type="deviance"))

#Lognormal
fit_AFT_loglog <- survreg(Surv(time, censor) ~ drug + age, data, dist='loglogistic')
fit_AFT_loglog %>% 
  summary()

fit_AFT_loglog %>% 
    AIC()

plot(residuals(fit_AFT_loglog, type="deviance"))

## Not much difference in terms of AIC between the models. No clear winner. I would default to Weibull or exponential.
## Need to look into this more.

## Going to try the same with the flexsurv package - Sameer suggested checking this out.

fit1 <- flexsurvreg(Surv(time, censor) ~ drug + age, data = data, dist = "weibull")
fit2 <- flexsurvreg(Surv(time, censor) ~ drug + age, data = data, dist = "exponential")
fit3 <- flexsurvreg(Surv(time, censor) ~ drug + age, data = data, dist = "lognormal")
fit4 <- flexsurvreg(Surv(time, censor) ~ drug + age, data = data, dist = "llogis")

## Getting the same results as before, except the direction of the exponential coefficient flips.
##Not sure what this means. I'm going to stick with the normal method.

##Let's check some Hazard functions
plot(fit1, type='cumhaz')
plot(fit2, type='cumhaz')
plot(fit3, type='cumhaz') ## Looks like Log-Norm or LogLog fits this the best! More Justification!!!
plot(fit4, type='cumhaz')

##Plot the Curves with the KM Estimates
##I'll plot each seperately and evaluate fit, then show the final one.

#Exp
plot(fit1, conf.int=T, lty = 1, xlab='Time (Months)', ylab='Survival')
legend('topright', c('KM', 'Exponential'), lwd=c(1,2), lty=1, col=c("black","red"))

plot(fit_KM_drug, conf.int=F, xlab='Time (Months)', ylab='Survival')
lines(predict(fit_AFT_exp, newdata=list(drug='0', age=36),type="quantile",p=seq(.01,.99,by=.01)),seq(.99,.01,by=-.01),col="#8E8E8E", lwd=2)
lines(predict(fit_AFT_exp, newdata=list(drug='1', age=36),type="quantile",p=seq(.01,.99,by=.01)),seq(.99,.01,by=-.01),col="#5CB2C8", lwd=2)
legend('topright', c('Drug 0', 'Drug 1'), lwd=2, lty=2, col=c("#8E8E8E","#5CB2C8"))

#Wei
plot(fit2, conf.int=F, lty = 1, xlab='Time (Months)', ylab='Survival', ci=F, col='#5CB2C8', lwd=2)
lines(survfit(coxfit2), conf.int=F, col='#8E8E8E', lwd=2)
legend('topright', c('KM','Cox-PH', 'Weibull'), lwd=c(2,2,2), lty=1, col=c('black',"#8E8E8E","#5CB2C8"))


plot(fit_KM_drug, conf.int=F, xlab='Time (Months)', ylab='Survival', lty=c(1,2), lwd=2)
lines(predict(fit_AFT_W, newdata=list(drug='0', age=36),type="quantile",p=seq(.01,.99,by=.01)),seq(.99,.01,by=-.01),col="#5CB2C8", lwd=2)
lines(predict(fit_AFT_W, newdata=list(drug='1', age=36),type="quantile",p=seq(.01,.99,by=.01)),seq(.99,.01,by=-.01),col="#5CB2C8", lwd=2, lty=2)

##Add the Cox predictions
lines(survfit(coxfit2, newdata=data.frame(drug='0', age=36)), conf.int=F, lwd=2, col="#8E8E8E", lty=1)
lines(survfit(coxfit2, newdata=data.frame(drug='1', age=36)), conf.int=F, lwd=2, col="#8E8E8E", lty=2)
legend('topright', c('KM', 'Cox-PH','Log-Normal'), lwd=c(2,2,2), lty=c(1,1,1), col=c('black' ,"#8E8E8E","#5CB2C8"))

text(20, 0.45, 'Drug 0')
text(5, 0.15, 'Drug 1')


#Log-Normal
plot(fit3, conf.int=F, lty = 1, xlab='Time (Months)', ylab='Survival', ci=F, col='#5CB2C8', lwd=2)
lines(survfit(coxfit2), conf.int=F, col='#8E8E8E', lwd=2)
legend('topright', c('KM','Cox-PH', 'Log-Normal'), lwd=c(2,2,2), lty=1, col=c('black',"#8E8E8E","#5CB2C8"))

plot(fit_KM_drug, conf.int=F, xlab='Time (Months)', ylab='Survival', lty=c(1,2), lwd=2)
lines(predict(fit_AFT_lognorm, newdata=list(drug='0', age=36),type="quantile",p=seq(.01,.99,by=.01)),seq(.99,.01,by=-.01),col="#5CB2C8", lwd=2)
lines(predict(fit_AFT_lognorm, newdata=list(drug='1', age=36),type="quantile",p=seq(.01,.99,by=.01)),seq(.99,.01,by=-.01),col="#5CB2C8", lwd=2, lty=2)

##Add the Cox predictions
lines(survfit(coxfit2, newdata=data.frame(drug='0', age=36)), conf.int=F, lwd=2, col="#8E8E8E", lty=1)
lines(survfit(coxfit2, newdata=data.frame(drug='1', age=36)), conf.int=F, lwd=2, col="#8E8E8E", lty=2)
legend('topright', c('KM', 'Cox-PH','Log-Normal'), lwd=c(2,2,2), lty=c(1,1,1), col=c('black' ,"#8E8E8E","#5CB2C8"))

text(20, 0.45, 'Drug 0')
text(5, 0.15, 'Drug 1')

#Log-Logistic
plot(fit3, conf.int=T, lty = 1, xlab='Time (Months)', ylab='Survival')
legend('topright', c('KM', 'Log-Logistic'), lwd=c(1,2), lty=1, col=c("black","red"))

plot(fit_KM_drug, conf.int=F, xlab='Time (Months)', ylab='Survival')
lines(predict(fit_AFT_loglog, newdata=list(drug='0', age=36),type="quantile",p=seq(.01,.99,by=.01)),seq(.99,.01,by=-.01),col="#8E8E8E", lwd=2)
lines(predict(fit_AFT_loglog, newdata=list(drug='1', age=36),type="quantile",p=seq(.01,.99,by=.01)),seq(.99,.01,by=-.01),col="#5CB2C8", lwd=2)
legend('topright', c('Drug 0', 'Drug 1'), lwd=2, lty=2, col=c("#8E8E8E","#5CB2C8"))

## Both methods show that at about 10 months, there will be 0% survival. 
## Similar steepness to the curve too.


## Survival of someone who is 50 at drug entry.
plot(fit3, newdata=data.frame(drug='0', age=50), col=c("#8E8E8E"))
lines(fit3, newdata=data.frame(drug='1', age=50), col=c("#5CB2C8"))

## Let's combine into a plot for reporting.

plot(predict(fit_AFT_lognorm, newdata=list(drug='0', age=50),type="quantile",p=seq(.01,.99,by=.01)),seq(.99,.01,by=-.01),col="#8E8E8E", lwd=2, type='l', xlab='Time (Months)', ylab='Survival')
lines(predict(fit_AFT_lognorm, newdata=list(drug='1', age=50),type="quantile",p=seq(.01,.99,by=.01)),seq(.99,.01,by=-.01),col="#8E8E8E", lwd=2)
lines(survfit(coxfit2, newdata=data.frame(drug='0', age=50)), conf.int=F, lwd=2, col="#5CB2C8", lty=1)
lines(survfit(coxfit2, newdata=data.frame(drug='1', age=50)), conf.int=F, lwd=2, col="#5CB2C8", lty=1)
legend('topright', c('Cox-PH', 'Log-Normal'), col=c('#5CB2C8','#8E8E8E'), lwd=2)

text(10, 0.25, 'Drug 0')
text(2, 0.15, 'Drug 1')

```
<p> Ideally, I would like to find some sort of plot that tests the log-normal assumption. But I have searched high and low and can't find anything. Visual inspection will just have to do, along with AIC.</p>
