---
title: "SPPH 581c - Assignment 4"
author: "Patrick Daniele"
date: "February 20, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<p>I'm going to complete this as an R-Markdown document as some of the packages I use require R-Studio and a very up to date version of R.</p>


<p>Load Libraries </p>
```{r cars}
library(survival)
library(tidyverse)
library(survminer)
library(ggfortify)
```
<p>Incase you haven't seen it before, the "Tidyverse" is a set of packages that makes using R much better.</p>


<p>Import the data and basic tidying</p>
```{r}
data <- read.csv('C:/Users/pdaniele/Desktop/SPPH 581c/Assignment 4/alzheimers.csv', header=T)
str(data)

#Note to self, dates are factors. I don't think we need them, so I won't do anything about them now.

#Change drug to a factor
data <- data %>% 
  mutate(drug=factor(drug))

# Change it to a tibble for easier processing
data <- as_data_frame(data)
data
```

<p> Three majors sections:</p>

<ol>
  <li>Non-Parametric: Kaplan Meier Estimates</li>
  <li>Semi-Parametric: Cox Proportional Hazards Model</li>
  <li>Fully Parametric: Weibull or Fully Distributional Model</li>
</ol>


<h1>Non-Parametric</h1>
```{r}
## Overall
fit_KM <- survfit(Surv(time, censor) ~ 1, data)

fit_KM %>% 
  summary()

## Try different survival plot methods
#Extensions of GGPlot2

fit_KM %>% 
  ggsurvplot(risk.table = TRUE, ggtheme=theme_classic())

fit_KM %>% 
  autoplot(conf.int=F)

## By Drug
fit_KM_drug <- survfit(Surv(time, censor) ~ drug, data)

fit_KM_drug %>% 
  summary()

## Try different survival plot methods
fit_KM_drug %>% 
  autoplot(conf.int=F)

fit_KM_drug %>% 
  ggsurvplot(risk.table = TRUE, ggtheme=theme_classic())
```

<h1>Semi-Parametric</h1>
```{r}
coxfit1 <- coxph(Surv(time, censor) ~ drug, data)
coxfit1 %>% 
  summary()

autoplot(survfit(coxfit1, newdata=data.frame(drug=1)), conf.int=F)

## Need to Assess the Proportional Hazards Assumption
coxfit1 %>% 
  cox.zph()

## Looks OK to me. Check Schoenfeld/Martingale Residuals next.
```

<h1>Fully-Parametric (Accelerated Failure Time Models)</h1>
```{r}
#Exponential Distribution
fit_AFT_exp <- survreg(Surv(time, censor) ~ as.factor(drug) + age, data, dist='exponential')
fit_AFT_exp %>% 
  summary()

#Weibull Distribution
fit_AFT_W <- survreg(Surv(time, censor) ~ as.factor(drug) + age, data, dist='weibull')
fit_AFT_W %>% 
  summary()

#Lognormal
fit_AFT_lognorm <- survreg(Surv(time, censor) ~ as.factor(drug) + age, data, dist='lognormal')
fit_AFT_lognorm %>% 
  summary()
```
