---
title: "SPPH 581c - Assignment 4"
author: "Patrick Daniele"
date: "February 20, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<p>I'm going to complete this as an R-Markdown document as some of the packages I use require R-Studio and a very up to date version of R.</p>


<p>Load Libraries </p>
```{r cars}
library(survival)
library(tidyverse)
library(survminer)
library(ggfortify)
library(flexsurv)
```
<p>Incase you haven't seen it before, the "tidyverse" is a set of packages that makes using R much better (In my opinion).</p>


<p>Import the data and basic tidying</p>
```{r}
data <- read.csv('C:/Users/pdaniele/Desktop/SPPH 581c/Assignment 4/alzheimers.csv', header=T)
str(data)

#Note to self, dates are factors. I don't think we need them, so I won't do anything about them now.

#Change drug to a factor
data <- data %>% 
  mutate(drug=factor(drug))

# Change it to a tibble for easier processing
data <- as_data_frame(data)
data
```


<p> Basic data exploration</p>
```{r}
## Checking normality of Age.
hist(data$age)
## Looks good to me

#Checking how many had events
table(data$censor)

## Checking How many on each drug
table(data$drug)
## Balanced!
```

<p> Three majors sections:</p>

<ol>
  <li>Non-Parametric: Kaplan Meier Estimates</li>
  <li>Semi-Parametric: Cox Proportional Hazards Model</li>
  <li>Fully Parametric: Weibull, Exponential, Log-Normal, Log-Logistic</li>
</ol>


<h1>Non-Parametric</h1>
```{r}
## Overall
fit_KM <- survfit(Surv(time, censor) ~ 1, data)

fit_KM %>% 
  summary()

## Try different survival plot methods
#Extensions of GGPlot2

fit_KM %>% 
  ggsurvplot(risk.table = TRUE, ggtheme=theme_classic())

fit_KM %>% 
  autoplot(conf.int=F)

## By Drug
fit_KM_drug <- survfit(Surv(time, censor) ~ drug, data)

fit_KM_drug %>% 
  summary()

survdiff(Surv(time, censor) ~ drug, data)

## Try different survival plot methods
fit_KM_drug %>% 
  autoplot(conf.int=F)

fit_KM_drug %>% 
  ggsurvplot(risk.table = TRUE, ggtheme=theme_classic())


## Let's get both the by drug and overall on the same plot
plot(fit_KM, conf.int=F, xlab='Time (Months)', ylab='Survival', lwd=2)
lines(fit_KM_drug, lty=c(2,2), lwd=c(2,2), col=c("#8E8E8E","#5CB2C8"))
legend('topright', c('Drug 0', 'Drug 1'), lwd=2, lty=2, col=c("#8E8E8E","#5CB2C8"))
text(40, 1, 'Log-rank p=0.000971')

```

<h1>Semi-Parametric</h1>
```{r}
## Without Age
coxfit1 <- coxph(Surv(time, censor) ~ drug, data)
coxfit1 %>% 
  summary()

## With Age
coxfit2 <- coxph(Surv(time, censor) ~ drug + age, data)
coxfit2 %>% 
  summary()

## Check model Fit and Assumptions (PH)
coxfit1 %>% 
    AIC()

coxfit1 %>% 
  cox.zph() %>% 
  plot()

coxfit2 %>% 
    AIC()

coxfit2 %>% 
  cox.zph() %>% 
  plot()

## Global test looks good. Plots look OK to me. Check Schoenfeld/Martingale Residuals next.
##Age improves fit.

## Plotting survival for the drug=1 group.

autoplot(survfit(coxfit1, newdata=data.frame(drug=1)), conf.int=F)

```

<h1>Fully-Parametric</h1>
```{r}
#Exponential Distribution
fit_AFT_exp <- survreg(Surv(time, censor) ~ drug + age, data, dist='exponential')
fit_AFT_exp %>% 
  summary()

fit_AFT_exp %>% 
    AIC()

#Weibull Distribution
fit_AFT_W <- survreg(Surv(time, censor) ~ drug + age, data, dist='weibull')
fit_AFT_W %>% 
  summary()

fit_AFT_W %>% 
    AIC()

#Lognormal
fit_AFT_lognorm <- survreg(Surv(time, censor) ~ drug + age, data, dist='lognormal')
fit_AFT_lognorm %>% 
  summary()

fit_AFT_lognorm %>% 
    AIC()

#Lognormal
fit_AFT_loglog <- survreg(Surv(time, censor) ~ drug + age, data, dist='loglogistic')
fit_AFT_loglog %>% 
  summary()

fit_AFT_loglog %>% 
    AIC()

## Not much difference in terms of AIC between the models. No clear winner. I would default to Weibull or exponential.
## Need to look into this more.

## Going to try the same with the flexsurv package - Sameer suggested checking this out.

flexsurvreg(Surv(time, censor) ~ drug + age, data = data, dist = "weibull")
flexsurvreg(Surv(time, censor) ~ drug + age, data = data, dist = "exponential")
flexsurvreg(Surv(time, censor) ~ drug + age, data = data, dist = "lognormal")
flexsurvreg(Surv(time, censor) ~ drug + age, data = data, dist = "llogis")

## Getting the same results as before, except the direction of the exponential coefficient flips.
##Not sure what this means. I'm going to stick with the normal method.

##Plot the Curves with the KM Estimates


## To do!

```

